---
title: "Bootcamp R: Analysis and Modelling - a Tidy way"
output:
  ioslides_presentation:
    widescreen: yes
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(imager)
myimg <- load.image("./tidy-1.png")

```
# R

## [what we are going to do:] An overview of R

### R is very powerful. Statistics roots, thus deep data analysis heritage.


## pros and cons
### Benefits from

- wide availability of packages
- diversity in implementation
- audited and tested by community
- good documentation
  
### Challenges (for a newbie)
- Does not follow regular computer science programming paradigms
- package and function discovery can seem overwhelming 
  - (same/similar methods implemented slightly differently)
- some packages might not play well with others right out of the box

## Tidy framework

#### The tidyverse is a collection of R packages for data analysis that are developed with common ideas and norms. 

#### From Wickham et al. (2019):*“At a high level, the tidyverse is a language for solving data science challenges with R code. Its primary goal is to facilitate a conversation between a human and a computer about data. Less abstractly, the tidyverse is a collection of R packages that share a high-level design philosophy and low-level grammar and data structures, so that learning one package makes it easier to learn the next.”*

- Design for Humans
- Reuse existing data structures
- Design for pipes and functional programming
  
  
## Tidy-ness of data:

### Principles or characteristics that make a dataset tidy:

1. Each variable must have its own column.
2. Each observation must have its own row.
3. Each value must have its own cell.
``` {r, echo = FALSE}
plot(myimg)

```

# Pipes %>%

##
    Lets work through an example of the childrens' poem 'Little Bunny Foo Foo'
      example often cited by Hadley Wickham as he described piping.
  
      Little bunny Foo Foo
      Went hopping through the forest
      Scooping up the field mice
      And bopping them on the head
      
## Scripting Steps

  As a code, we create a 'little_bunny' object called 'foo_foo'

    foo_foo <- little_bunny()

  Foo foo performs a bunch of actions like hop, scoop and bop

    foo_foo_1 <- hop(foo_foo, through = forest)
    foo_foo_2 <- scoop(foo_foo_1, up = field_mice)
    foo_foo_3 <- bop(foo_foo_2, on = head)
  
  We end up creating a bunch of intermediate objects to hold values

##  Functional way:
  we can write the steps as a functional composition

    bop(
      scoop(
        hop(foo_foo, through = forest),
        up = field_mice
       ), 
     on = head
    )
    
  
## Using Pipes
    The code becomes a lot more human readable, and thus easier to maintain and debug.
    
      `foo_foo %>%
      hop(through = forest) %>%
      scoop(up = field_mice) %>%
      bop(on = head)`

  This is a good way to approach modelling and data analysis but will not be ideal or would not work for some situations.

## Tidy Package universe
    list a table of packages by tidy-verse vs outside
    1. Data import
    2. Data cleaning and wrangling
    3. EDA
    4. Modelling
    5. Visualisation
    6. Reporting, Collaboration
    
## Exercises for today
    1. Rework the Machine Learning 1 - Python bootcamp example in R, using Tidy Models
        a. PCA (& UMP)
        b. Linear Regression (Random Forest and LM)
    2. Discuss Tidyverse and Tidy Model
    3. Discuss and work in Rmarkdown
    3. Tidy Models vs Caret vs ML3
    
    
## References:
      Websites
      - Statistical Tools for High-throughput Data Analysis (STHDA)
        <http://www.sthda.com/english/>
      - Tidy Models
        <https://www.tidymodels.org>
      - Learning R
        <https://education.rstudio.com/learn/>
      
      Books:
      - R for Data Science
        <https://r4ds.had.co.nz/tidy-data.html>
      - R Markdown Cookbook
        <https://bookdown.org/yihui/rmarkdown-cookbook/>
      - Tidy Modeling with R
        <https://www.tmwr.org/>
      - Feature Engineering
        <http://www.feat.engineering/>
